Modeling site-specific amino-acid preferences deepens phylogenetic estimates of viral divergence

### 2018-05-09 manuscript/manuscript.tex
updated references

More recent work has used mutation-selection models to better account for purifying selection~\citep{halpern1998evolutionary,yang2008mutation,rodrigue2010mutation,tamuri2012estimating,mccandlish2014modeling}.
These models explicitly incorporate the fact that different protein sites prefer different amino acids, and so can improve phylogenetic estimates when there are deep branches~\citep{philippe1998good, lartillot2007suppression, le2008phylogenetic, si2008empirical, wang2008class}.
However, these approaches require inferring the site-specific purifying selection from natural sequence data.

@@

\subsection*{pbMutSel inference with \texttt{PhyloBayes-MPI}.}
For \ref{fig:phylobayes}, we fit a pbMutSel model to the full HA tree. 
We ran one chain for 5500 steps, saved every sample, and discarded the first 550 samples as a burnin. 
We used \texttt{PhyloBayes-MPI} program \texttt{readpb\_mpi} to compute the majority-rule consensus tree and the posterior average site-specific amino-acid preferences. 
Convergence was assessed visually using Tracer (\url{http://tree.bio.ed.ac.uk/software/tracer/}) and by the correlation of amino-acid preferences inferred by two independent chains (r=0.996).

@@

\subsection*{Acknowledgments}
We thank Erick Matsen and Trevor Bedford for helpful comments about the project and manuscript.
SKH is supported in part by training grant T32AI083203 from the NIAID of the National Institutes of Health.
This work was supported by the NIAID and NIGMS of the NIH under grant numbers R01AI127893 and R01GM102198. 
JDB is supported in part by a Faculty Scholars grant from the Howard Hughes Medical Institute and the Simons Foundation.
The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
All code and data used for the analyses in this paper is available at https://github.com/jbloomlab/ divergence_timing_manuscript.

>>>>

More recent work has used mutation-selection models to better account for purifying selection~\citep{halpern1998evolutionary,yang2008mutation,rodrigue2010mutation,tamuri2012estimating,mccandlish2014modeling}.
These models explicitly incorporate the fact that different protein sites prefer different amino acids, and so can improve phylogenetic estimates when there are deep branches~\citep{philippe1998good, lartillot2007suppression, le2008phylogenetic, si2008empirical, wang2008class, susko2018accelerated}.
However, these approaches require inferring the site-specific purifying selection from natural sequence data.

@@

\subsection*{pbMutSel inference with \texttt{PhyloBayes-MPI}.}
For \ref{fig:phylobayes}, we fit a pbMutSel model to the full HA tree. 
We ran one chain for 5500 steps, saved every sample, and discarded the first 550 samples as a burnin. 
We used \texttt{PhyloBayes-MPI} program \texttt{readpb\_mpi} to compute the majority-rule consensus tree and the posterior average site-specific amino-acid preferences. 
Convergence was assessed visually using Tracer \citep{rambaut2018posterior} and by the correlation of amino-acid preferences inferred by two independent chains (r=0.996).

@@

\subsection*{Acknowledgments}
We thank Erick Matsen and Trevor Bedford for helpful comments about the project and manuscript.
SKH is supported in part by training grant T32AI083203 from the NIAID of the National Institutes of Health.
This work was supported by the NIAID and NIGMS of the NIH under grant numbers R01AI127893 and R01GM102198. 
JDB is supported in part by a Faculty Scholars grant from the Howard Hughes Medical Institute and the Simons Foundation.
The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
All code and data used for the analyses in this paper is available at \url{https://github.com/jbloomlab/ divergence_timing_manuscript}.
